Alleviating Topology Mismatch in Distributed Overlay Networks: A Survey

### 2014-01-27 unstructured.tex
2012.12.23 (Mema) Text reworked.

\section{Decentralized Unstructured Architectures}
\label{section:unstructured}

This section is dedicated to the unstructured P2P architectures. In the next
paragraphs, algorithms for tackling the topology mismatch problem on such
networks are discussed, analyzed and categorized based on the use of the
overlay structure, message forwarding approach for peer communication and
techniques proposed for detecting proximity in order to optimize overlay
topology.

@@

The directed BFS%, as shown in Figure~\ref{figure:dbfs:dbfs},
tries to avoid this delay by forwarding the query messages only to a selected
set of neighbors, in which the selection criteria varies from the number of
results received previously, distance in terms of hops, bandwidth, or the query
load of the neighbor.

In local indices%, as shown in Figure~\ref{figure:dbfs:localindx},
each node
indexes data within a radius of $r$ hops and uses this local index to answer
queries on behalf of them without generating additional traffic. Local indices
greatly reduce the aggregate bandwidth usage of the network and improves query
efficiency, however, updating the indices in cases with frequent node joins and
leaves introduces a serious overhead to the system if the radius is kept broad.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{ \bf Delay Aware P2P System}
A new \emph{Delay Aware P2P System (DAPS)} is introduced in \cite{ZL2005}. Its
main goal is to reduce the time of a look-up request by dividing the routing
tables of peers into several sectors of increasing delay. The source node
emitting the query message defines a delay boundary or the \emph{pruning
factor} in the paper's context. Request messages are then forwarded only to
nodes whose delay is less than or equal that boundary. With the clustered
routing tables and the loose organization of DAPS' overlay network it is
considered by is between structured and unstructured.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Distributed Cycle Minimization Protocol (DCMP)}
We have already discussed that blind flooding mechanism broadcasts messages with
no real targeting, many times to a direction far from the path that can
ultimately fulfill the query request. Another problem with topology unaware
systems is the duplication of messages due to cycles, even along the correct
forwarding path. The authors of \cite{ZKB2008} focused on that exact deficiency
of overlay networks and introduced Distributed Cycle Minimization Protocol
(DCMP), a dynamic, fully distributed method to remove cycles without sacrificing
overlay connectivity, resilience and other important properties of unstructured
architectures by avoiding a hierarchical organization of peers. In DCMP, once a
cycle is detected, the most powerful node in that cycle is elected as the
\emph{Gate Peer} and the cycle is then broken at a strategic place so that it
will result in the minimization of the distance between the Gate Peer and all
other nodes that participated in that cycle. The process is managed by using two
special message types namely \emph{Information Collection (IC)} and \emph{Cut
Message (CM)}. One disadvantage of DCMP is that since the distance a forwarded
message can travel is limited by the TTL value, which is practically $7$ in most
cases DCMP cannot detect cycles formed by more than 7 peers. Even though cycle
elimination improves the network performance, it does not actually solve the
topology mismatch problem.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Replication Strategies in Unstructured Peer-to-Peer Networks}
\cite{CS2002} aim to improve the inefficient blind search algorithm by
replicating the data in a P2P network. The main intuition behind the
idea of replication, or using cached copies, is that as the number of copies for
each item increases in the network, it would be easier for a search algorithm,
even a random one, to locate these items. In order to analyze the feasibility of
such an approach, the authors investigate three different replication
strategies, namely uniform, proportional, and square root allocation. In the
uniform model, the copies of items are uniformly replicated in the network,
while in the proportional model, the items are replicated based on their query
rate, so that frequently queried items are replicated more. The square root
allocation is a strategy proposed by Cohen and Shenker, which is a model between
the uniform and proportional allocation.  In order to evaluate the outcome of
the replication models, the overall costs of successful and unsuccessful
searches in the network are compared. Results argue that the uniform allocation
model minimizes the maximum search size, therefore reduces the time spent on an
unsuccessful search. The proportional model, on the other hand, promotes the
more frequently queried items by replicating them more, therefore decreases the
search time for popular items, but suffers when needing to locate the rare
items. The authors also claim that the expected successful search size is the
same for uniform and proportional models, and any approach between them would
always behave much better. Therefore they propose the square root allocation
approach, which is a replication strategy that minimizes the expected search
size of successful queries in P2P networks.

@@

% \begin{figure}[ht]
% \centering
% \subfigure[Underlying network with edge costs.] {
%   \includegraphics[scale=0.4]{img/algorithms/narada}
%   \label{figure:narada:underlying}
% }\qquad\qquad
% \subfigure[Peer A sends a message to the rest using regular broadcast.] {
%   \includegraphics[scale=0.4]{img/algorithms/narada2}
%   \label{figure:narada:regu}
% }\qquad\qquad
% \subfigure[Peer A sends a message to the rest using end system multi-cast.] {
%   \includegraphics[scale=0.4]{img/algorithms/narada3}
%   \label{figure:narada:multicast}
% }
% \caption{A visualization of the Narada protocol.}
% \label{figure:narada}
% \end{figure}

\emph{Narada} \cite{CRZ2000,CRSZ2001,CRSZ2002} is a generic
protocol for creating self-adapting overlay networks, that can achieve
application-layer multi-cast communication without requiring IP multi-cast
infrastructure at the network layer\footnote{
  IP multi-cast is the term referring to the method of sending IP datagrams to a
  group of receivers in a single transmission. Even though the method is
  available some years now, IP multi-cast has not taken off as anticipated since
  many believe that it violates the stateless design of the current Internet.
  For this reason it is widely deployed in contained environments like
  enterprises (i.e. IPTV applications - distance learning, video conferencing),
  commercial stock exchanges, and multimedia content delivery networks.
}. Although it was not originally designed as a P2P system nor can be considered
scalable in the contemporary sense of the term, Narada has been pioneering
(maybe along with \emph{Scattercast} \cite{C2000}) in that
it was the first to consider the feasibility of overlay-based, application
layer, services over the Internet that can take into account bandwidth and
latency properties of the physical underlying infrastructure. Additionally,
authors realized the inefficiency caused by the topology mismatch problem and
tried to encounter it by building a richer connected graph, called a
\emph{mesh}, and building per source minimum spanning trees\footnote{
  ENarada \cite{LYL2008} used Gossip protocol for the
construction
}.
%as shown in Figure~\ref{figure:narada}.
They also keep the graph and the trees
dynamically updated while nodes continue to join and leave the network.
Moreover, the protocol tries to ease the physical link stress, the overall
resource usage and the relative delay among end systems. Unfortunately, the most
important limitation of Narada is that although it works reasonably well for
small groups, it does not scale well for larger networks, therefore it is not
suitable for, potentially very large, P2P file sharing application networks.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Adaptive Overlay Topology Optimization (AOTO)}
\emph{Adaptive Overlay Topology Optimization (AOTO)} \cite{LZXN2003} is one
of the first attempts, along with Narada, to address the topology mismatch
problem. AOTO is a distributed algorithm that works in two steps, namely
\emph{Selective Flooding} and \emph{Active Topology} to optimize the usage of
the underlying physical resources. In \emph{Selective Flooding} a minimum
spanning tree is built among each peer and its immediate neighbors so that
queries are not flooding the whole network and at the same time without
shrinking the search scope. During \emph{Active Topology}, the overlay
links are revised so that they depict the physical network topology as close as
possible. This is done independently, by each peer by replacing non-flooding
neighbors, with closer nodes. Picking a replacement out of these non-flooding
neighbors follows a random policy (called \emph{Randomized AT} algorithm in the
paper's context). For these steps a peer needs to keep track of communication
costs to all its neighbors (e.g. the network delay) as well as between any
pair of neighbors (meaning also that additional, cost probing, message types
need to be added to the Gnutella protocol). The randomized AT algorithm is
applied by the source peer everytime the neighbour list is updated or an updated
neighbour cost table is received.

@@

\begin{center}
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
%
medium &
% tracking of communication costs to all peer's neighbors as well as between
% any pair of neighbors. Whenever a new neighbor cost table is received or
% there is a change of neighbors, the source peer has to re-calculate the
% multi-cast tree and apply the randomized AT algorithm.
high &
% needs global knowledge
medium
\end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Adaptive Connection Establishment (ACE)}
\emph{Adaptive Connection Establishment (ACE)} \cite{LZXN2004} uses the network
delay as a metric to calculate the costs between nodes in a spanning tree.
Each peer calculates the costs to its logical one-hop
neighbors and forms a \emph{neighbor cost table (NCT)} using a special
routing message type. Two neighboring peers exchange their NCTs in order for
every peer to obtain the cost between any pair of its local neighbors forming a
small overlay topology. Moreover, based on obtained NCTs a minimum spanning tree
among each peer and its one-hop neighbors is built. Ultimately, physically
far away neighbors are replaced by physically close ones. Specifically, a peer
$S$ probes the distance between one of its non-flooding neighbor's neighbor,
say $G$ and $H$ respectively. Assuming that the neighbor's neighbor ($SH$) is
measured as being ``closer'' that the neighbor ($SG$), the link to the later
is cut and a new connection is established. In the opposite situation where
distance $SG$ is closer than $SH$ then if $SH$ is shorter that $GH$ then $H$ is
preserved as $S$'s new neighbor. In the case where $SH$ is larger than $SG$ and
$GH$, no connection will be established and $S$ will continue probing another
neighbor's neighbor. The above optimization is conducted within $1$-neighbor
closure (among its source peer and all its direct neighbors) but the scope can
be extended. The larger the scope, the better topology matching improvement but
also the greater the computational overhead.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Location-aware Topology Matching (LTM)}
\emph{Location-aware Topology Matching (LTM)} \cite{LLXNZ2004} proposes a
method to optimize the overlay structure of the P2P network based on the
physical topology. To achieve this, peers issue special messages called
\textit{TTL-detector}s with TTL values of 0 and 1 so that the receiving peers
discover one- and two-hop neighbor sets ($N$ and $N^2$ respectively) around
them and calculating communication costs. Calculation is achieved through
time-stamp checking, so clocks must be synchronized. The latency information
gathered is later used to evolve the overlay network into a more efficient one,
without reducing the search scope. Each node examines the latency information
among his direct neighbors and those with larger latency are put in to a
will-cut list where they remain for a certain period of time until when they are
ultimately cut and recorded to the peer's cut list. Thus, low productive
connections are dropped and replaced by more efficient ones, reducing the
latency on the overall network. Although LTM improves the overall efficiency of
the P2P network, it does not actually use real physical topology information,
just latency metrics, therefore it does not offer a guaranteed safety net for
the topology mismatch problem.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Scalable Bipartite Overlay (SBO)}
\emph{Scalable Bipartite Overlay (SBO)}
\cite{LXN2004,LXN2007} reduces the overhead of creating
and maintaining a minimum spanning tree cost by randomly dividing the nodes into
two groups, namely red and white and assigning different tasks to different
groups. When a peer joins the network, it is randomly assigned with an initial
color, say red or white (separating all peers into two groups). Then the peer
that plays the role of the boot-strap host provides the joining peer with a list
of active peers along with their color information. The later uses this list to
establish connections to differently colored peers. This way, all peers form a
bipartite overlay. The white peers measure distances to neighbors (red only) by
using the network delay as a metric and report their corresponding red
neighbors. The red peers, equipped with the information of all two-hop
neighbors ($N^2$), creates minimum spanning tree of these neighbors and
assigns efficient forwarding paths. Having a minimum spanning tree with two
hops, a red peer is able to send its queries within that range. Some white
peers, though, have become non-forwarding neighbors. In this phase, such a
(white) neighbor will try to find another red peer being two hops away from its
current red neighbor to replace the later as its new neighbor. The white peers
can further optimize their positions within the overlay if this is required.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Two-Hop-Away Neighbor Comparison and Selection (THANCS)}
Work in \cite{LNXE2005,L2008} proposes a distributed heuristic
called \emph{Two-Hop-Away Neighbor Comparison and Selection (THANCS)} that
tries to minimize the overlay hop costs. THANCS is considered a \emph{local
search method}, in the sense that it targets in finding a locally optimum
solution, by exploiting knowledge within a 2-hop radius. The algorithm consists
of two main components: \emph{piggybacking neighbor distance on queries} and
\emph{neighbor comparison and selection} which are further discussed bellow.
The first component dictates peers to probe distances (using network delay
distance measuring) with its immediate neighbors and stores information
locally. This is done by introducing a special query message type, the
\emph{Piggy Message (PM)} which includes information about the neighbor
identification and measured distance. A peer $p$ constructs a PM for its
neighbor $n$, which contains $n$'s IP address and the distance between them.
When $p$ receives a query from $n$, this PM will be piggybacked to the query
message that will be then forwarded to all other neighbors of peer $p$. 
Each neighbor will detach the PM information (it will not be further forwarded),
record the $pq$ distance and process the query as always after receiving a query
message. Selection of which incoming queries should piggyback a PM is proposed
by either \emph{pure probability-based (PPB)} or \emph{new neighbor triggered
(NNT)} policies. The second component, namely neighbor comparison and selection,
dictates peer $p$ to probe the distance to all his known, not yet probed,
two-hop neighbors ($ N^2(p)$). The distance of $pn$ is known to $p$. Upon
receiving a PM from node $n$ with the distance of $nq$, where $q$ is a direct
neighbor of $n$, $p$ chooses one of the following two paths. First, if $q$ is a
direct neighbor of $p$, then the later will check the cost of the involved
links. If the most cost-intensive link is one of $pq$ or $pn$, then it is put
into a \emph{will-cut} list. If it is $nq$ then nothing is done (since either
$n$ or $q$ will have the chance to handle it themselves).  On the other hand, if
$q$ is within two-hop radius from $p$, the later probes the former (if it hasn't
got the distance information yet) and again compares distances $pq$, $pn$ and
$nq$. If $pq$ is the most cost-intensive, $p$ will not establish the connection.
If it is $pn$, $p$ will establish connection $pq$ and put $pn$ in the will-cut
list. If it is $nq$, $p$ will stay connected with both $n$ and $q$,
such that eventually in later stages either $n$ or $q$ will disconnect link $nq$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TODO: DOUBLE CHECK IF THIS ALGORITHM GOES IN THIS SECTION/SUBSECTION
\subsubsection{Adaptive Peer Selection (APS)}
Bernstein et al., contributed on the use of machine learning for building peer
selection strategies from past experience \cite{BFLZ2003}. A decision tree is
used to rate peers based on information about connection characteristics that is
collected. These can be load, bandwidth, and past uploading experience. Then a
by Markov decision process is incorporated as a mechanism to shape the policy
for switching among the peers. The problem with this approach is that the peer
selection algorithm is slow due to the learning process and the complexity of
the method.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Innocuous Topology Aware (ITA) Overlay Construction}
\cite{PRFM2009} introduces an algorithm for Innocuous Topology Aware
construction of unstructured P2P networks. The paper's proposition is twofold.
Overlay construction and search method. For the overlay construction it employs
the notion of \emph{short} and \emph{long} connections. Assuming $N$ is the
number of all network participants, $\alpha \leq 1 $ a system-wide magic number
and $x$ an $\alpha$-related latency threshold, the boot-strapping peer randomly
selects $\alpha \ast N$ close (latency bellow $x$) and
$\left( 1 - \alpha \right) \ast N$ distant (latency above $x$) nodes as its
neighbors. The search method has two steps. First, the querying node initiates a
flood to its distant neighborhood with $TTL = 1$. Then, peers that receive a
query over a long link, start a local flood with $TTL = ttl$, were $ttl$ is a
system defined parameter.

The goal of the overlay construction is to create a network that exposes low
clustering, a beneficial characteristic of random graphs which can lead to a
larger coverage of peers with the same number of messages and reduced
duplication. This means more efficient information lookup that exposes low or no
negative impact whatsoever to the rest of the mechanisms employed in
unstructured P2P systems\footnote{For example the 1-hop replication and the
dynamic querying mechanisms of later versions of Gnutella.}. The paper reports
50\% reduction in communication latency among peers by cutting off some 20\% of
the IP-level message traffic generation.

@@

The topology adaptation based protocols modify the topology of the P2P network
using various techniques. The two commonly used approaches that can
back such feature and were investigated in this section are creating
\emph{spanning tree}s using connection graphs, and creating \emph{clusters} of
physically close nodes.

>>>>

\section{Unstructured P2P Networks}
\label{section:unstructured}

This section is dedicated to unstructured P2P networks. Algorithms
for tackling the topology mismatch problem in such
networks are discussed, analyzed, and categorized based on the use of the
overlay structure, the message forwarding approach for peer communication, and
techniques proposed for detecting proximity in order to optimize overlay
topology.

@@

The directed BFS
tries to avoid this delay by forwarding the query messages only to a select
set of neighbors, in which the selection criteria varies from the number of
results received previously, distance in terms of hops, bandwidth, or the query
load of the neighbor.

In local indices
each node
indexes data within a radius of $r$ hops and uses this local index to answer
queries on behalf of them without generating additional traffic. Local indices
greatly reduce the aggregate bandwidth usage of the network and improves query
efficiency, however, updating the indices in cases with frequent node joins and
leaves introduces a serious overhead into the system if the radius is kept broad.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{ \bf Delay Aware P2P System}
A new \emph{Delay Aware P2P System (DAPS)} is introduced in \cite{ZL2005}. Its
main goal is to reduce the time of a look-up request by dividing the routing
tables of peers into several sectors of increasing delay. The source node
emitting the query message defines a delay boundary or the \emph{pruning
factor} in the paper's context. Request messages are then forwarded only to
nodes whose delay is less than or equal that boundary. 
%With the clustered
%routing tables and the loose organization of DAPS' overlay network it is
%considered by is between structured and unstructured.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Distributed Cycle Minimization Protocol (DCMP)}
We have already discussed that blind flooding often causes messages to be routed in a direction
that is far from the path that can
ultimately fulfill the query request. Another problem with topology unaware
systems is the duplication of messages due to cycles, even along the correct
forwarding path. The authors of \cite{ZKB2008} focused on that exact deficiency
of overlay networks and introduced Distributed Cycle Minimization Protocol
(DCMP), a dynamic, fully distributed method to remove cycles without sacrificing
overlay connectivity, resilience and other important properties of unstructured
architectures by avoiding a hierarchical organization of peers. In DCMP, once a
cycle is detected, the most powerful node in that cycle is elected as the
\emph{Gate Peer} and the cycle is then broken at a strategic place so that it
will result in the minimization of the distance between the Gate Peer and all
other nodes that participated in that cycle. The process is managed by using two
special message types namely \emph{Information Collection (IC)} and \emph{Cut
Message (CM)}. One disadvantage of DCMP is that since the distance a forwarded
message can travel is limited by the TTL value, which is practically $7$ in most
cases, DCMP cannot detect cycles formed by more than 7 peers. Even though cycle
elimination improves the network performance, it does not actually solve the
topology mismatch problem.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Replication Strategies in Unstructured Peer-to-Peer Networks}
\cite{CS2002} aim to improve the inefficient blind search algorithm by
replicating the data in a P2P network. The main intuition behind the
idea of replication, or using cached copies, is that as the number of copies for
each item increases in the network, it would be easier for a search algorithm,
even a random one, to locate these items. To analyze the feasibility of
such an approach, the authors investigate three different replication
strategies, namely uniform, proportional, and square root allocation. In the
uniform model, the copies of items are uniformly replicated in the network,
while in the proportional model, the items are replicated based on their query
rate, so that frequently queried items are replicated more. The square root
allocation is a strategy proposed by Cohen and Shenker, which is a model between
the uniform and proportional allocation.  To evaluate the outcome of
the replication models, the overall costs of successful and unsuccessful
searches in the network are compared. Results argue that the uniform allocation
model minimizes the maximum search size, therefore reduces the time spent on an
unsuccessful search. The proportional model, on the other hand, promotes the
more frequently queried items by replicating them more and therefore decreases the
search time for popular items, but suffers when needing to locate the rare
items. The authors also claim that the expected successful search size is the
same for uniform and proportional models, and any approach between them would
always behave much better. Finally, they propose the square root allocation
approach, which is a replication strategy that minimizes the expected search
size of successful queries in P2P networks.

@@

% \begin{figure}[ht]
% \centering
% \subfigure[Underlying network with edge costs.] {
%   \includegraphics[scale=0.4]{img/algorithms/narada}
%   \label{figure:narada:underlying}
% }\qquad\qquad
% \subfigure[Peer A sends a message to the rest using regular broadcast.] {
%   \includegraphics[scale=0.4]{img/algorithms/narada2}
%   \label{figure:narada:regu}
% }\qquad\qquad
% \subfigure[Peer A sends a message to the rest using end system multicast.] {
%   \includegraphics[scale=0.4]{img/algorithms/narada3}
%   \label{figure:narada:multicast}
% }
% \caption{A visualization of the Narada protocol.}
% \label{figure:narada}
% \end{figure}

\emph{Narada} \cite{CRZ2000,CRSZ2001,CRSZ2002} is a generic
protocol for creating self-adapting overlay networks, that can achieve
application-layer multicast communication without requiring IP multicast
infrastructure at the network layer\footnote{
  IP multicast is the term referring to the method of sending IP datagrams to a
  group of receivers in a single transmission. Even though the method is
  available some years now, IP multicast has not taken off as anticipated since
  many believe that it violates the stateless design of the current Internet.
  For this reason, it is widely deployed in contained environments like
  enterprises (i.e. IPTV applications - distance learning, video conferencing),
  commercial stock exchanges, and multimedia content delivery networks.
}. Although it was not originally designed as a P2P system, Narada is
a pioneer in that 
%(and the accompanying \emph{Scattercast} \cite{C2000})
it was the first to consider the feasibility of overlay-based, application
layer, services over the Internet that can take into account bandwidth and
latency properties of the physical underlying infrastructure. Additionally,
authors realized the inefficiency caused by the topology mismatch problem and
tried to address it by building a richer connected graph, called a
\emph{mesh}, and building per source minimum spanning trees\footnote{
  ENarada \cite{LYL2008} used Gossip protocol for the
construction
}.
%as shown in Figure~\ref{figure:narada}.
They also keep the graph and the trees
dynamically updated while nodes continue to join and leave the network.
The protocol aims to ease the physical link stress, the overall
resource usage and the relative delay among end systems. Unfortunately, the most
important limitation of Narada is that although it works reasonably well for
small groups, it does not scale well for larger networks, therefore it is not
suitable for, potentially very large, P2P file sharing application networks.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Adaptive Overlay Topology Optimization (AOTO)}
\emph{Adaptive Overlay Topology Optimization (AOTO)} \cite{LZXN2003} is one
of the first attempts, along with Narada, to address the topology mismatch
problem. AOTO is a distributed algorithm that works in two steps, namely
\emph{Selective Flooding} and \emph{Active Topology} to optimize the usage of
the underlying physical resources. In \emph{Selective Flooding}, a minimum
spanning tree is built among each peer and its immediate neighbors so that
queries are not flooding the whole network and at the same time without
shrinking the search scope. During \emph{Active Topology}, the overlay
links are revised so that they depict the physical network topology as close as
possible. This is done independently, by each peer by replacing non-flooding
neighbors, with closer nodes. Picking a replacement out of these non-flooding
neighbors follows a random policy (called \emph{Randomized AT} algorithm in the
paper's context). For these steps a peer needs to keep track of communication
costs to all its neighbors (e.g. the network delay) as well as between any
pair of neighbors (meaning also that additional, cost probing, message types
need to be added to the Gnutella protocol). The randomized AT algorithm is
applied by the source peer every time the neighbour list is updated or an updated
neighbour cost table is received.

@@

\begin{center}
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
%
medium &
% tracking of communication costs to all peer's neighbors as well as between
% any pair of neighbors. Whenever a new neighbor cost table is received or
% there is a change of neighbors, the source peer has to re-calculate the
% multicast tree and apply the randomized AT algorithm.
high &
% needs global knowledge
medium
\end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Adaptive Connection Establishment (ACE)}
\emph{Adaptive Connection Establishment (ACE)} \cite{LZXN2004} uses the network
delay as a metric to calculate the costs between nodes in a spanning tree.
Each peer calculates the costs to its logical one-hop
neighbors and forms a \emph{neighbor cost table (NCT)} using a special
routing message type. Two neighboring peers exchange their NCTs in order for
every peer to obtain the cost between any pair of its local neighbors forming a
small overlay topology. Moreover, based on obtained NCTs a minimum spanning tree
among each peer and its one-hop neighbors is built. Ultimately, physically
far away neighbors are replaced by physically close neighbors. Specifically, a peer
$S$ probes the distance between one of its non-flooding neighbor's neighbors,
say $G$ and $H$ respectively. Assuming that the neighbor's neighbor ($SH$) is
measured as being ``closer'' than the neighbor ($SG$), the link to the later
is cut and a new connection is established. In the opposite situation where
distance $SG$ is closer than $SH$ then if $SH$ is shorter than $GH$, then $H$ is
preserved as $S$'s new neighbor. In the case where $SH$ is larger than $SG$ and
$GH$, no connection will be established and $S$ will continue probing another
neighbor's neighbor. The above optimization is conducted within $1$-neighbor
closure (among its source peer and all its direct neighbors) but the scope can
be extended. The larger the scope, the better topology matching improvement but
also the greater the computational overhead.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Location-aware Topology Matching (LTM)}
\emph{Location-aware Topology Matching (LTM)} \cite{LLXNZ2004} proposes a
method to optimize the overlay structure of the P2P network based on the
physical topology. To achieve this, peers issue special messages called
\textit{TTL-detector}s with TTL values of 0 and 1 so that the receiving peers
discover one- and two-hop neighbor sets ($N$ and $N^2$ respectively) around
them and calculating communication costs. Calculation is achieved through
timestamp checking, so clocks must be synchronized. The latency information
gathered is later used to evolve the overlay network into a more efficient one,
without reducing the search scope. Each node examines the latency information
among his direct neighbors and those with larger latency are placed on a
will-cut list where they remain for a certain period of time after which they are
cut and placed on the peer's cut list. Thus, low productive
connections are dropped and replaced by more efficient ones, reducing the
latency on the overall network. Although LTM improves the overall efficiency of
the P2P network, it does not actually use real physical topology information,
just latency metrics, therefore it does not offer a guaranteed safety net for
the topology mismatch problem.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Scalable Bipartite Overlay (SBO)}
\emph{Scalable Bipartite Overlay (SBO)}
\cite{LXN2004,LXN2007} reduces the overhead of creating
and maintaining a minimum spanning tree cost by randomly dividing the nodes into
two groups, namely red and white and assigning different tasks to different
groups. When a peer joins the network, it is randomly assigned with an initial
color, say red or white (separating all peers into two groups). Then the peer
that plays the role of the bootstrap host provides the joining peer with a list
of active peers along with their color information. The later uses this list to
establish connections to differently colored peers. This way, all peers form a
bipartite overlay. The white peers measure distances to neighbors (red only) by
using the network delay as a metric and report their corresponding red
neighbors. The red peers, equipped with the information of all two-hop
neighbors ($N^2$), create minimum spanning tree of these neighbors and
assigns efficient forwarding paths. Having a minimum spanning tree with two
hops, a red peer is able to send its queries within that range. Some white
peers, though, have become non-forwarding neighbors. In this phase, such a
(white) neighbor will try to find another red peer being two hops away from its
current red neighbor to replace the later as its new neighbor. The white peers
can further optimize their positions within the overlay if this is required.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Two-Hop-Away Neighbor Comparison and Selection (THANCS)}
Work in \cite{LNXE2005,L2008} proposes a distributed heuristic
called \emph{Two-Hop-Away Neighbor Comparison and Selection (THANCS)} that
tries to minimize the overlay hop costs. THANCS is considered a \emph{local
search method} because it aims to find a locally optimum
solution by exploiting knowledge within a 2-hop radius. The algorithm consists
of two main components: \emph{piggybacking neighbor distance on queries} and
\emph{neighbor comparison and selection} which are further discussed below.
The first component requires peers to probe distances (using network delay
distance measuring) with its immediate neighbors and stores information
locally. This is done by introducing a special query message type, the
\emph{Piggy Message (PM)} which includes information about the neighbor
identification and measured distance. A peer $p$ constructs a PM for its
neighbor $n$, which contains $n$'s IP address and the distance between them.
When $p$ receives a query from $n$, this PM will be piggybacked to the query
message that will be then forwarded to all other neighbors of peer $p$. 
Each neighbor will detach the PM information (it will not be further forwarded),
record the $pq$ distance and process the query as always after receiving a query
message. Selection of which incoming queries should piggyback a PM is proposed
by either \emph{pure probability-based (PPB)} or \emph{new neighbor triggered
(NNT)} policies. The second component, namely neighbor comparison and selection,
dictates peer $p$ to probe the distance to all his known, not yet probed,
two-hop neighbors ($ N^2(p)$). The distance of $pn$ is known to $p$. Upon
receiving a PM from node $n$ with the distance of $nq$, where $q$ is a direct
neighbor of $n$, $p$ chooses one of the following two paths. First, if $q$ is a
direct neighbor of $p$, then the latter will check the cost of the involved
links. If the most cost-intensive link is one of $pq$ or $pn$, then it is put
into a \emph{will-cut} list. If it is $nq$ then nothing is done (since either
$n$ or $q$ will have the chance to handle it themselves).  On the other hand, if
$q$ is within two-hop radius from $p$, the later probes the former (if it hasn't
got the distance information yet) and again compares distances $pq$, $pn$, and
$nq$. If $pq$ is the most cost-intensive, $p$ will not establish the connection.
If it is $pn$, $p$ will establish connection $pq$ and put $pn$ in the will-cut
list. If it is $nq$, $p$ will stay connected with both $n$ and $q$,
such that eventually in later stages either $n$ or $q$ will disconnect link $nq$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TODO: DOUBLE CHECK IF THIS ALGORITHM GOES IN THIS SECTION/SUBSECTION
\subsubsection{Adaptive Peer Selection (APS)}
Bernstein et al., use machine learning techniques for building peer
selection strategies from past experience \cite{BFLZ2003}. A decision tree is
used to rate peers based on information about connection characteristics that is
collected. These can be load, bandwidth, and past uploading experience. Then a
Markov decision process is incorporated as a mechanism to shape the policy
for switching among the peers. Although novel, the downside of this approach is that the peer
selection algorithm is slow due to the learning process and the complexity of
the method.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Innocuous Topology Aware (ITA) Overlay Construction}
\cite{PRFM2009} introduces an algorithm for Innocuous Topology Aware
construction of unstructured P2P networks. The paper's proposition is twofold.
Overlay construction and search method. For the overlay construction, it employs
the notion of \emph{short} and \emph{long} connections. Assuming $N$ is the
number of all network participants, $\alpha \leq 1 $ a system-wide magic number
and $x$ an $\alpha$-related latency threshold, the bootstrapping peer randomly
selects $\alpha \ast N$ close (latency below $x$) and
$\left( 1 - \alpha \right) \ast N$ distant (latency above $x$) nodes as its
neighbors. The search method has two steps. First, the querying node initiates a
flood to its distant neighborhood with $TTL = 1$. Then, peers that receive a
query over a long link start a local flood with $TTL = ttl$, where $ttl$ is a
system defined parameter.

The goal of the overlay construction is to create a network that exhibits low
clustering, a beneficial characteristic of random graphs which can lead to a
larger coverage of peers with the same number of messages and reduced
duplication. This means more efficient information lookup that has minimal or no
negative impact whatsoever to the rest of the mechanisms employed in
unstructured P2P systems\footnote{For example the 1-hop replication and the
dynamic querying mechanisms of later versions of Gnutella.}. The paper reports
50\% reduction in communication latency among peers by cutting off some 20\% of
the IP-level message traffic generation.

@@

The topology adaptation based protocols modify the topology of the P2P network
using various techniques. The two most commonly used approaches to 
topology adaptation are creating
\emph{spanning tree}s using connection graphs, and creating \emph{clusters} of
physically close nodes.
